---
title: "wordle_analysis"
output: pdf_document
date: "2025-08-14"
---

```{r setup, include=FALSE}
library(jsonlite)
```

## Preamble
This is a template for extracting and graphing important statistics and doing comparison/multiple test.


### Claude
```{r}
claude_control <- fromJSON("predictions/claude_control.json")
claude_control_guesses = claude_control$guess_results$num_guesses 
print(mean(claude_control_guesses))
print(var(claude_control_guesses))


claude_experiment <- fromJSON("predictions/claude_hypothesis.json")
claude_experiment_guesses = claude_experiment$guess_results$num_guesses 
print(mean(claude_experiment_guesses))
print(var(claude_experiment_guesses))
```

H0: experiment set is no higher than control set
HA: experiment set is higher

Catch: Cannot assume normal distribution of data, see graph for reason

```{r}
hist(claude_control_guesses, breaks = 2.5:7.5)
```
```{r}
hist(claude_experiment_guesses, breaks = 2.5:7.5)
```

Hypothesis Testing:
```{r}
wilcox.test(claude_experiment_guesses, claude_control_guesses, alternative = "greater", exact = TRUE)

# Still need to do Permutation Test here.

```

### GPT
```{r}
gpt_control <- fromJSON("predictions/gpt_control.json")
gpt_control_guesses = gpt_control$guess_results$num_guesses 
print(mean(gpt_control_guesses))
print(var(gpt_control_guesses))


gpt_experiment <- fromJSON("predictions/gpt_hypothesis.json")
gpt_experiment_guesses = gpt_experiment$guess_results$num_guesses 
print(mean(gpt_experiment_guesses))
print(var(gpt_experiment_guesses))
```

H0: experiment set is no higher than control set
HA: experiment set is higher

Catch: Cannot assume normal distribution of data, see graph for reason

```{r}
hist(gpt_control_guesses, breaks = 1.5:7.5)
```
```{r}
hist(gpt_experiment_guesses, breaks = 3.5:7.5)
```

Hypothesis Testing:
```{r}
wilcox.test(gpt_experiment_guesses, gpt_control_guesses, alternative = "greater", exact = TRUE)

# Still need to do Permutation Test here.
```


### Mistral
```{r}
mistral_control <- fromJSON("predictions/mistral_control.json")
mistral_control_guesses = mistral_control$guess_results$num_guesses 
print(mean(mistral_control_guesses))
print(var(mistral_control_guesses))


mistral_experiment <- fromJSON("predictions/mistral_hypothesis.json")
mistral_experiment_guesses = mistral_experiment$guess_results$num_guesses 
print(mean(mistral_experiment_guesses))
print(var(mistral_experiment_guesses))
```

H0: experiment set is no higher than control set
HA: experiment set is higher

Catch: Cannot assume normal distribution of data, see graph for reason

```{r}
hist(mistral_control_guesses, breaks = 2.5:7.5)
```
```{r}
hist(mistral_experiment_guesses, breaks = 3.5:7.5)
```

Hypothesis Testing:
```{r}
wilcox.test(mistral_experiment_guesses, mistral_control_guesses, alternative = "greater", exact = TRUE)

# Still need to do Permutation Test here.
```

### Discuss
#### Why do we use not use a t-test?
Because data is not normal. Does it look like normal to you?

#### And what tests do we use?
Wilcoxon and Permutation. Explained here for why it is appropriate.



